---
title: "class05: data vis with ggplot"
author: "Sneha Patel"
format: pdf
toc: true 
---

# Our first plot

R base graphics
```{r}
plot(cars)
```
 must install and load ggplot2 library before you can use it 
 install.packages("ggplot2")
```{r}
library(ggplot2) 
```

```{r}
ggplot(cars)
```
 ^ wont work
 
 every ggplot needs at least 3 layers
 
 - **data** (ie the data.frame we have)
 
 - **aes** (the aesthetic mapping of our data to what we want to plot) 
 
 - **geoms** (how we want to plot this stuff)

```{r}
ggplot(data=cars) +
  aes(x=speed, y=dist) +
  geom_point() +
  geom_smooth(method=lm, se=FALSE)
```
# A very cool plot
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```
```{r}
p <- ggplot(data=genes) + aes(x=Condition1, y=Condition2, col=State) + geom_point(alpha=0.4) 
p
```

```{r}
up_table <- table(genes$State=="up")

up_table[2]
```

There are `r nrow(genes)` genes in this data set.

There are `r ncol(genes)` columns in this data set, with titles `r colnames(genes)`

Here are how many genes are upregulated: `r up_table[2]`

Fraction of total genes upregulated: `r up_table[2] / nrow(genes) * 100`

```{r}
p + scale_colour_manual( values=c("pink","gray","turquoise") )
```
p + labs(title="Gene response to Drugs", x = "drug", y = "control"

```{r}
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)

library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)

ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.4)
```

```{r}
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, color = pop, size = pop) +
  geom_point(alpha=0.8) + scale_size_area(max_size=10)
```
```{r}
gapminder_1957 <- gapminder %>% filter(year==1957)

ggplot(gapminder_1957) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.7) + scale_size_area(max_size=15) 
```
```{r}
gapminder_1957_2007 <- gapminder %>% filter(year==1957 | year==2007)

ggplot(gapminder_1957_2007) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.7) + scale_size_area(max_size=15) + (facet_wrap(~year))
```
```{r}
gapminder_top5 <- gapminder %>% 
  filter(year==2007) %>% 
  arrange(desc(pop)) %>% 
  top_n(5, pop)

ggplot(gapminder_top5) + 
  aes(x = reorder(country,-lifeExp), y = lifeExp, fill=gdpPercap) + labs(title="Life expectancy of the 5 biggest countries") + geom_col()
```
```{r}
ggplot(mtcars, aes(mpg, disp)) + geom_point()
```
```{r}
ggplot(mtcars) + aes(x=mpg, y=disp) + geom_point()
```

