---
title: "class10: halloween candy mini project"
author: "Sneha Patel"
format: pdf
---

# Background

In this mini-project we will examine 538 Halloween Candy data.

First step is read the data
```{r}
candy <- read.csv("candy-data.csv", row.names=1)
head(candy)
```
```{r}
nrow(candy)
```

# Q1. 
85 types of candy in the dataset

```{r}
table(candy$fruity)
```

# Q2
There are 38 fruity candy types

```{r}
rownames(candy)
```

```{r}
candy["HersheyÕs Special Dark", ]$winpercent
```
```{r}
candy["Kit Kat", ]$winpercent
```
```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

# Q3
My favorite candy is Hershey's Special Dark and the win percent is 59.2%

# Q4 
Kit Kat's win percent is 76.77%

# Q5 
Tootsie Roll Snack Bars win percent is 49.65%

```{r}
skimr::skim(candy)
```
# Q6
winpercent seems to be on a different scale

# Q7 
The 1s and 0s in the chocolate column represent true or false aka if it is chocolate or not

# Q8

```{r}
hist(candy$winpercent)
```
```{r}
library(ggplot2)

ggplot(candy) + aes (x = winpercent) + geom_histogram(bins=15, col="black", fill="lightblue")
```
# Q9 
The distribution of `winpercent` is not symmetrical

```{r}
median(candy$winpercent)
```

# Q10 
The center of the distribution is below 50%

```{r}
#choc_avg <- mean(candy$winpercent(chocolate==1))
choc_true <-  as.logical(candy$chocolate)
choc_all <- candy[choc_true,]$winpercent
choc_avg <- mean(choc_all)

fruity_true <-  as.logical(candy$fruity)
fruity_all <-candy[fruity_true,]$winpercent
fruity_avg <- mean(fruity_all)

choc_avg > fruity_avg
```

# Q11
Chocolate candies are ranked higher than fruit candy

```{r}
t.test(choc_all, fruity_all)
```

# Q12
The difference is not statistically significant 

```{r}
library(dplyr)
candy %>% arrange(winpercent) %>% head(5)
```
# Q13
The 5 least liked candies are Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters

```{r}
candy %>% arrange(winpercent) %>% tail(5)
```

# Q14
The top 5 candies are Snickers, Kit Kat, Twix, Reese's Miniatures, and Reese's Peanut Butter cup

```{r}
library(ggplot2)

ggplot(candy) + aes(winpercent, rownames(candy)) + geom_col()
```
```{r}
library(ggplot2)

ggplot(candy) + aes(winpercent, reorder(rownames(candy), winpercent)) + geom_col()
```
```{r}
ggsave("mybarplot.png")
```
# Adding color
Let’s setup a color vector (that signifies candy type) that we can then use for some future plots. We start by making a vector of all black values (one for each candy). Then we overwrite chocolate (for chocolate candy), brown (for candy bars) and red (for fruity candy) values.
```{r}
my_cols <- rep("pink", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "brown"
my_cols[as.logical(candy$bar)] = "lightblue"
my_cols[as.logical(candy$fruity)] = "orange"
```

```{r}
library(ggplot2)

ggplot(candy) + aes(winpercent, reorder(rownames(candy), winpercent)) + geom_col(fill=my_cols)
```
# Q17 
Worst ranked chocolate is sixlets

# Q18
Best ranked fruity candy is starburst

# Taking a look at pricepercent
What about value for money? What is the the best candy for the least money? One way to get at this would be to make a plot of winpercent vs the pricepercent variable. The pricepercent variable records the percentile rank of the candy’s price against all the other candies in the dataset. Lower vales are less expensive and high values more expensive.

To this plot we will add text labels so we can more easily identify a given candy. There is a regular geom_label() that comes with ggplot2. However, as there are quite a few candys in our dataset lots of these labels will be overlapping and hard to read. To help with this we can use the geom_text_repel() function from the ggrepel package.

```{r}
library(ggrepel)

ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)

```

```{r}
#highest winpercent, lowest pricepercent
ord <- order(candy$pricepercent, decreasing = FALSE)
head(candy[ord,c(11,12)], n=5 )
```

# Q19
Tootsie roll midgies is the biggest bang for your buck

```{r}
#lowest winpercent, highest pricepercent
ord <- order(candy$pricepercent, decreasing = TRUE)
head(candy[ord,c(11,12)], n=5 )
```
# Q20
Worst expensive candy is Nik L Nip

# Exploring the correlation structure
```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

# Q22
Fruit and Chocolate are anti-correlated

# Q23 
Chocolate and bar/winpercent are the most positively correlated

# PCA
```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```
```{r}
plot(pca$x[,1], pca$x[,2], col=my_cols, pch=16)
```

```{r}
my_data <- cbind(candy, pca$x[,1:3])

p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)
p
```
```{r}
p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")

```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
# Q24
Fruity, hard, and pluribus have strong PC1 in the positive direction. This makes sense since chocolate isn't hard, and hard candies typically come in individually wrapped bite size pieces
